Dependent Object Types

### 2012-08-10 article.tex
almost there

\subsubsection{Set of Values}
$\relv k \Gamma s T$ defines the set of values that appear to have
type $T$ when taking at most $k$ steps. $\Gamma$ and $s$ must agree:
$\dom(\Gamma) = \dom(s)$ (ordered) and $\forall (x : T) \in \Gamma, x
\in \relv k \Gamma s T$. A variable $y$ belongs to $\relv 0 \Gamma s
T$ simply by being in the store. In addition, it belongs to $\relv k
\Gamma s T$ for $k > 0$, if it defines all method and value labels in
the expansion of $T$ appropriately for $j < k$ steps.
\begin{align*}
&\relv k \Gamma s T = \{ y \ |\\
&\gap \Gamma \ts T \wfe \andl\\
&\gap y \in \dom(s) \andl\\
&\gap \forall j < k,\\
&\gap y \mapsto T_c \ldefs{\seq{l = v}\;\seq{m(x)=t}} \in s,\\
&\gap \Gamma \ts T \expand_y \seq{D},\\
&\gap (\forall m_i : S \rightarrow U \in \seq{D},\\
&\gap\gap \Gamma \ts S \wf \andl t_i \in \rele j {\Gamma \envplus{x : S}} s U)\\
&\gap \andl\\
&\gap (\forall l_i : V \in \seq{D},\\
&\gap\gap v_i \in \relv j \Gamma s V)\\
&\}
\end{align*}

This relation captures the observation that the only ways for a term
to get stuck is to have a field selection on an uninitialized field or
a method invocation on uninitialized method.

Note that this relation doesn't directly impose constraints on the
type labels. In particular, since we're just enforcing appropriate
initializations of all fields and methods by looking at an expansion,
lower bounds of type selections are completely ignored. This is by
design, because a term doesn't get stuck in a method invocation if
its argument type doesn't satisfy the lower bound required by a type
selection -- the upper bound is what matters operationally.

However, a {\it potential pitfall} is that the value itself might
occur in the types $S$, $U$, $V$, because we substitute it for the
``self'' occurrences in the expansion.

@@

\begin{proofsketch}
Because $S$ is a subtype of $U$, it should hold that the expansion of
$S$ subsumes the expansion of $U$. Therefore, for $v \in \relv k
\Gamma s U$, we have fewer definitions to check than for $v \in \relv
k \Gamma s S$.

There are two {\it potential pitfalls}.

First, because we don't check for type labels, we can end up in a
situation where a type label defined in both $S$ and $U$ is actually
only used meaningfully in $U$, and so $v \in \relv k \Gamma s S$ never
enforces it. Consider $v$ to map to $\Top \refine z {\ldecl l \Top}
\ldefs{l = v}$, $S$ is $\Top \refine z {\Ldecl L \Top \Top , \ldecl l
  \Top}$ and $U$ is $\Top \refine z {\Ldecl L \Top \Top , \ldecl l
  {z.\Top}}$. Now, $v \in \relv k \Gamma s S$ even though it doesn't
define the type label $L$. But $v \not\in \relv k \Gamma s U$ because
$v.L$ is not well-formed. \remark{Counterexample!}

Another worry is potential narrowing issues for methods: because the
parameter type is contravariant, $U$ can require a more precise type
than $S$. This should be fine, because if the parameter type becomes
non-expanding, then the check is vacuously true.
\end{proofsketch}

>>>>

\subsubsection{Set of Values}
$\relv k \Gamma s T$ defines the set of values that appear to have
type $T$ when taking at most $k$ steps. $\Gamma$ and $s$ must agree:
$\dom(\Gamma) = \dom(s)$ (ordered) and $\forall (x : T) \in \Gamma, x
\in \relv k \Gamma s T$. A variable $y$ belongs to $\relv 0 \Gamma s
T$ simply by being in the store. In addition, it belongs to $\relv k
\Gamma s T$ for $k > 0$, if it defines all type, method and value labels in
the expansion of $T$ appropriately for $j < k$ steps.
\begin{align*}
&\relv k \Gamma s T = \{ y \ |\\
&\gap \Gamma \ts T \wfe \andl\\
&\gap y \in \dom(s) \andl\\
&\gap \forall j < k,\\
&\gap y \mapsto T_c \ldefs{\seq{l = v}\;\seq{m(x)=t}} \in s,\\
&\gap \Gamma \ts T \expand_y \seq{D},\\
&\gap (\forall L_i : S \rightarrow U \in \seq{D},\\
&\gap\gap \Gamma \ts v.L_i \wfe \spcomma S \wf \spcomma U \wf \spcomma S \sub U\\
&\gap (\forall m_i : S \rightarrow U \in \seq{D},\\
&\gap\gap \Gamma \ts S \wf \spcomma U \wf \andl t_i \in \rele j {\Gamma \envplus{x : S}} s U)\\
&\gap \andl\\
&\gap (\forall l_i : V \in \seq{D},\\
&\gap\gap \Gamma \ts V \wf \andl v_i \in \relv j \Gamma s V)\\
&\}
\end{align*}

This relation captures the observation that the only ways for a term
to get stuck is to have a field selection on an uninitialized field or
a method invocation on uninitialized method. However, a {\it potential
  pitfall} is that the value itself might occur in the types $S$, $U$,
$V$, because we substitute it for the ``self'' occurrences in the
expansion, so the relation makes sure that the type labels exist.

@@

\begin{proofsketch}
Because $S$ is a subtype of $U$, it should hold that the expansion of
$S$ subsumes the expansion of $U$, when the ``self'' occurrences are
of type $S$. Therefore, for $v \in \relv k \Gamma s U$, we have fewer
declarations to check than for $v \in \relv k \Gamma s S$.

A {\it potential pitfall} is whether some types of the expansion of
$U$ can become non-expanding when the ``self'' occurrences are
actually $v$ instead of just abstractly of type $S$, causing a check
to fail. Another worry is that such a non-expanding type results from
narrowing of a parameter type. However, we believe this is fine by
subtyping regularity wtr to $\nswfe$ and $v \in \relv k \Gamma s S$
which cheks bounds and $\nswfe$ of the type labels.
\end{proofsketch}
